---
title: "Smoke Outlook for `r params$BURN_NAME` Prescribed Fire"
date: "Issued at `r Sys.time()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
geometry: margin=0.25in
params: 
  BURN_NAME: ""
  CONTACT_INFO: ""
  LON: ""
  LAT: ""
  MODEL_RUN: ""
  BURN_DATE: ""
  RUN_ID_URL: ""
  MD_PATH: ""
  MON_RADIUS: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## SPATIAL
library(sp)
library(rgeos)
library(raster)
library(rgdal)
library(maptools)
library(ggmap)
library(sf)
library(png)

## DATA MANAGEMENT
library(tidyverse)
library(skimr)
library(readxl)
# library(zoo)
library(curl)
library(RCurl)
library(rjson)

## PLOTTING
library(scales)
library(units)
library(viridis)
library(extrafont)
library(gtable)
library(grid)
library(ggimage)
library(gridExtra)
library(knitr)
library(kableExtra)
library(egg)
library(tidycensus)
library(ggrepel)
library(leaflet)
library(htmlwidgets)


## SMOKE DATA
library(PWFSLSmoke)
library(MazamaSpatialUtils)
```


```{r burn_info, include = FALSE}
#############################################################################
## pull monitors near burn
#############################################################################

##***************************************
## ENTER BURN INFORMATION HERE
##***************************************

burn_info <- tibble(lon = params$LON, 
                    lat = params$LAT,
                    name = params$BURN_NAME)

#------

## Pull airnow data and reshapre for plotting

# number of previous hours to pull monitoring data
n_hrs <- 96

# pull hourly data within 100 km of rx burn  and the last n_days hours
airnow <- airnow_loadLatest() %>% 
  monitor_subsetByDistance(burn_info$lon, burn_info$lat, radius = params$MON_RADIUS) 

airnow$data <- airnow$data %>% 
  slice(-(1:(n()-n_hrs)))


# get number of monitors 
n_mons <- length(unique(airnow$meta$monitorID))


# calculate mean and prep data for plotting
mon_dailyMean_list <- monitor_dailyStatistic(airnow, FUN = get("mean"), 
                                        dayStart = "midnight", 
                                        na.rm = TRUE, 
                                        minHours = 18)

mon_names <- mon_dailyMean_list[[1]]$siteName

# calculate mean daily AQI for preceeding days
mon_mean_df <- mon_dailyMean_list[[2]] 
colnames(mon_mean_df)[2:(2+length(mon_names)-1)] <- mon_names
mon_mean_df <- mon_mean_df %>% 
  gather(mon_name, pm_val, -1) %>% 
  mutate(aqi_col = aqiColors(pm_val),
         aqi_cat = cut(pm_val, 
                       breaks = AQI$breaks_24,
                       labels = AQI$names))

# calculate most recent 24hr average AQI
rec_daily_df <- mon_mean_df %>% 
  group_by(mon_name) %>% 
  arrange(desc(datetime)) %>% 
  slice(1) %>% 
  ungroup() %>% 
  rename(siteName = mon_name) %>% 
  left_join(., airnow$meta, by = "siteName") 


# read aqi table to include in final output
aqi_tab <- read_csv("../raw_data/aqi_table.csv")

# air specialist contact info
# air_spec_contact <- tibble(name = c("Jeremy Ash", "Ralph Perron", "Trent Wickman"),
#                            number = c("608-234-3300", "603-536-6228", "218-626-4372"),
#                            email = c("jeremy.ash@usda.gov", "ralph.perron@usda.gov", "trent.wickman@usda.gov")) %>% 
#   filter(name == params$AUTHOR)

# air_spec_contact <- params$AUTHOR

```




This report documents the recent air quality, weather forecast and smoke analysis for the `r burn_info$name` prescribed fire, planned for `r params$BURN_DATE`. For more information: `r params$CONTACT_INFO`. 



## Recent Air Quality in Area 

Shown below is the recently measured air quality index (AQI) in the area. The map shows the location of the burn and the most recent daily AQI for nearby permanent monitors. The figure shows the daily AQI for the previous 3 days for the monitors in the map.

```{r burn_map_aqi, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# set up google API for ggmap 
api <- readLines("../gis/google.api")
register_google(key = api)

# read in burn icon for mapping
burn_icon <- tibble(x = burn_info$lon,
                        y = burn_info$lat,
                        image = "../icons/redFlame.png")

##***************************************
## PULLING BURN MAP INFO: MAY NEED TO ADJUST ZOOM
##***************************************

# pull burn map from google
burn_map_info <- get_map(location = c(burn_info$lon, burn_info$lat),
                    maptype = "terrain",
                    zoom = 8)

# plot burn map
burn_map <- ggmap(burn_map_info, extent = "device") +
  geom_image(aes(x = x, y = y, image = image), data = burn_icon) +
  geom_point(aes(x = longitude, y = latitude, color = aqi_cat), size = 3, show.legend = FALSE, data = rec_daily_df) +
  geom_label_repel(aes(x = longitude, y = latitude, label=siteName),
                   box.padding = 1,
                   arrow = arrow(length = unit(0.02, "npc")), 
                   data = rec_daily_df) +
  scale_color_manual(values = rec_daily_df$aqi_col)


# recent AQI values at nearby monitors
daily_aqi_plot <- ggplot(aes(x = datetime, y = pm_val, group = mon_name), data = mon_mean_df) +
  geom_bar(aes(fill = aqi_cat), stat = "identity", show.legend = FALSE) +
  scale_fill_manual(values = AQI$colors) +
  facet_wrap(~mon_name, ncol = 1) +
  labs(y = "AQI", x = NULL) +
  theme_minimal()


# aqi_table 
aqi_kable <- aqi_tab %>%
  mutate('Air Quality Index (AQI)' = cell_spec(c("Good", "Moderate", "USG", "Unhealthy", "Very Unhealthy", "Hazardous"), color = "black", bold = T, background = AQI$colors)) %>% 
  kable(escape = FALSE, linesep = "") %>%
  kable_styling(full_width = F, position = "left")


# show the burn map and aqi side-by-side
grid.arrange(burn_map, daily_aqi_plot, nrow = 1, widths = c(2:1))

# show aqi color scheme and explanations
aqi_kable

```

**Note**: Generally, smoke sensitive people are kids and older adults along with those who have cardiovascular or respiratory conditions.

For more smoke and health information: [https://airnow.gov/](https://airnow.gov/)


## Modeled Smoke Trajectories
Below are the BlueSky modeling outputs for the day of the planned burn. The interactive map below shows the predicted average smoke conditions and maximum smoke conditions for `r params$BURN_DATE`. Select either "Daily Average" or "Daily Maximum" in the map and zoom in/out to change the viewing extent. The average smoke conditions are helpful for concerns related to public health. The maximum smoke conditions show the highest concentration at each point on the map for that day. This map is useful for identifying potential impacts to visibility on roadways or other areas.

__Smoke conditions on `r params$BURN_DATE`__
```{r leaflet_map_data_mgt, include = FALSE, echo = FALSE, message = FALSE, warning = FALSE}

##-------------
## general data
##-------------

# bluesky PM color codes and binds
bsky_col <- read_csv("../raw_data/bsky_gray_color_ramp.csv",
                     trim_ws = TRUE) 



##-------------
## download and extract files
##-------------

# smoke md directory
md_smoke <- paste("../md/",
                  params$MD_PATH,
                  "/",
                  params$MODEL_RUN, sep = "")


# smoke dispersion
smoke_disp_link <- paste(params$RUN_ID_URL,
                         "/output/smoke_dispersion.kmz", sep = "")

# download smoke dispersion into a zip and regular version 
curl_download(smoke_disp_link, destfile = paste(md_smoke,
                                                "/smoke_dispersion.kmz.zip",
                                                sep = ""))

curl_download(smoke_disp_link, destfile = paste(md_smoke,
                                                "/",
                                                str_replace_all(Sys.Date(), "-", ""),
                                                "_",
                                                gsub(" ", "_", params$BURN_NAME),
                                                "_",
                                                params$MODEL_RUN, 
                                                ".kmz", sep = ""))

# create the directory to store files
dir.create(paste(md_smoke,
                 "/smoke_dispersion_files", sep = ""))

# unzip the file
unzip(zipfile = paste(md_smoke,
                      "/smoke_dispersion.kmz.zip",
                      sep = ""),
      exdir = paste(md_smoke,
                 "/smoke_dispersion_files", sep = ""))

# get rid of zip file
file.remove(paste(md_smoke,
                      "/smoke_dispersion.kmz.zip",
                      sep = ""))

# grid info to get extent of png files
grid_info_link <- paste(params$RUN_ID_URL, 
                        "/output/grid_info.json", sep = "")

# download grid info
curl_download(grid_info_link, destfile = paste(md_smoke, 
                                               "/smoke_dispersion_files/grid_info.json", sep = ""))


# burn location info and UTC offset
burn_loc_link <- paste(params$RUN_ID_URL, 
                       "/output/data/fire_locations.csv", sep = "")

# download burn locations
curl_download(burn_loc_link, destfile = paste(md_smoke,
                                              "/smoke_dispersion_files/burn_locations.csv", sep = ""))
  


##-------------
## re-code rasters
##-------------

# burn ctds
burn_loc <- read_csv(paste(md_smoke,
                           "/smoke_dispersion_files/burn_locations.csv", sep = ""))

# burn CRS
burn_crs <- st_crs(read_sf(paste(md_smoke,
                                 "/smoke_dispersion_files/doc.kml", sep = "")))

# read PNG files
day_one_max <- readPNG(paste(md_smoke,
                             "/smoke_dispersion_files/100m_daily_maximum_", 
                              burn_loc$date_time, 
                              "_UTC",
                              gsub(":", "", burn_loc$utc_offset), 
                             ".png", sep = ""))

day_one_avg <- readPNG(paste(md_smoke, 
                             "/smoke_dispersion_files/100m_daily_average_", 
                             burn_loc$date_time, 
                             "_UTC",
                             gsub(":", "", burn_loc$utc_offset),
                             ".png", sep = ""))


# pull extent from grid_info.json
grid_info <- fromJSON(file=paste(md_smoke,
                                 "/smoke_dispersion_files/grid_info.json", sep = ""))[[1]]


# # read in burn icon for mapping
# burn_icon <- tibble(x = burn_loc$longitude,
#                     y = burn_loc$latitude,
#                     image = "../icons/redFlame.png")


# collapse RGB bands into one with hex codes as values
day_one_max_r <- as.raster(day_one_max)
day_one_avg_r <- as.raster(day_one_avg)

# replace hex codes with PM categories
day_one_max_r[day_one_max_r == "#C8C8C8B2"] <- bsky_col$pm[bsky_col$hex == "#C8C8C8B2"] #1
day_one_max_r[day_one_max_r == "#AFAFAFB2"] <- bsky_col$pm[bsky_col$hex == "#AFAFAFB2"] #2
day_one_max_r[day_one_max_r == "#969696B2"] <- bsky_col$pm[bsky_col$hex == "#969696B2"] #3
day_one_max_r[day_one_max_r == "#7D7D7DB2"] <- bsky_col$pm[bsky_col$hex == "#7D7D7DB2"] #4
day_one_max_r[day_one_max_r == "#646464B2"] <- bsky_col$pm[bsky_col$hex == "#646464B2"] #5
day_one_max_r[day_one_max_r == "#4B4B4BB2"] <- bsky_col$pm[bsky_col$hex == "#4B4B4BB2"] #6
day_one_max_r[day_one_max_r == "#323232B2"] <- bsky_col$pm[bsky_col$hex == "#323232B2"] #7
day_one_max_r[day_one_max_r == "#191919B2"] <- bsky_col$pm[bsky_col$hex == "#191919B2"] #8
day_one_max_r[day_one_max_r == "#00000000"] <- bsky_col$pm[bsky_col$hex == "#00000000"] #0

day_one_avg_r[day_one_avg_r == "#C8C8C8B2"] <- bsky_col$pm[bsky_col$hex == "#C8C8C8B2"] #1
day_one_avg_r[day_one_avg_r == "#AFAFAFB2"] <- bsky_col$pm[bsky_col$hex == "#AFAFAFB2"] #2
day_one_avg_r[day_one_avg_r == "#969696B2"] <- bsky_col$pm[bsky_col$hex == "#969696B2"] #3
day_one_avg_r[day_one_avg_r == "#7D7D7DB2"] <- bsky_col$pm[bsky_col$hex == "#7D7D7DB2"] #4
day_one_avg_r[day_one_avg_r == "#646464B2"] <- bsky_col$pm[bsky_col$hex == "#646464B2"] #5
day_one_avg_r[day_one_avg_r == "#4B4B4BB2"] <- bsky_col$pm[bsky_col$hex == "#4B4B4BB2"] #6
day_one_avg_r[day_one_avg_r == "#323232B2"] <- bsky_col$pm[bsky_col$hex == "#323232B2"] #7
day_one_avg_r[day_one_avg_r == "#191919B2"] <- bsky_col$pm[bsky_col$hex == "#191919B2"] #8
day_one_avg_r[day_one_avg_r == "#00000000"] <- bsky_col$pm[bsky_col$hex == "#00000000"] #0


# convert to matrix and change values to numeric
day_one_max_m <- matrix(as.numeric(day_one_max_r),
                        nrow = dim(day_one_max_r)[1],
                        ncol = dim(day_one_max_r)[2],
                        byrow = TRUE)

day_one_avg_m <- matrix(as.numeric(day_one_avg_r),
                        nrow = dim(day_one_avg_r)[1],
                        ncol = dim(day_one_avg_r)[2],
                        byrow = TRUE)


# convert to raster
day_one_max_ras <- raster(day_one_max_m)
day_one_avg_ras <- raster(day_one_avg_m)


# ratify raster (convert to factor)
day_one_max_ras <- ratify(day_one_max_ras)
day_one_avg_ras <- ratify(day_one_avg_ras)

# add levels of factor in dataframe
levels(day_one_max_ras) <- data.frame(ID = c("1", "2", "3", "4", "5", "6", "7", "8"),
                                      pm = c("1-12", "12-35", "35-55", "55-150", "150-250", "250-350", "350-500", ">500"))


levels(day_one_avg_ras) <- data.frame(ID = c("1", "2", "3", "4", "5", "6", "7", "8"),
                                      pm = c("1-12", "12-35", "35-55", "55-150", "150-250", "250-350", "350-500", ">500"))


# set extent of raster
extent(day_one_max_ras) <- extent(c(grid_info[1], 
                                    grid_info[3],
                                    grid_info[2],
                                    grid_info[4]))

extent(day_one_avg_ras) <- extent(c(grid_info[1], 
                                    grid_info[3],
                                    grid_info[2],
                                    grid_info[4]))

# same crs as the doc.kml
crs(day_one_max_ras) <- burn_crs[["proj4string"]]
crs(day_one_avg_ras) <- burn_crs[["proj4string"]]

# list of two rasters
day_one_ls <- list(day_one_max_ras, day_one_avg_ras)
names(day_one_ls) <- c("day_one_max_ras", "day_one_avg_ras")
# return(day_one_ls)


# day_one_ls <- bsky_raster_gen_fun("20200124")


##-------------
## plot rasters
##-------------

# pull out rastersand burn_icon into separate files
day_one_max_ras <- day_one_ls[["day_one_max_ras"]]
day_one_avg_ras <- day_one_ls[["day_one_avg_ras"]]


# bluesky color ramp
pm_col_ramp <- bsky_col %>% 
  pull(hex_2)
pm_col_ramp <- pm_col_ramp[-9]

# create color palette function
pal <- colorFactor(palette = pm_col_ramp,
                   levels = levels(day_one_max_ras)[[1]][1]$ID,
                   # levels = levels(day_one_max_ras)[[1]][2]$pm,
                   # domain = NULL,
                   na.color = "transparent")

# generate leaflet map
day_one_leaflet <- leaflet(width = "100%") %>% 
  addTiles() %>%
  addRasterImage(day_one_max_ras, 
                 colors = pal,
                 opacity = 0.8, 
                 group = "Daily Maximum") %>%
  addRasterImage(day_one_avg_ras, 
                 colors = pal,
                 opacity = 0.8, 
                 group = "Daily Average") %>%
  addLegend(pal = pal, 
            # values = day_one_max_ras,
            # values = levels(day_one_max_ras)[[1]][2]$pm,
            values = levels(day_one_max_ras)[[1]][1]$ID,
            # labels =  c("1-12", "12-55", "35-55", "55-150", "150-250", "250-350", "350-500", ">500"),
            labFormat  = labelFormat(
              transform = function(x) {
                levels(day_one_max_ras)[[1]]$pm[which(levels(day_one_max_ras)[[1]]$ID == x)]}),
            opacity = 0.8, 
            title = "PM2.5 (ug m^-3)") %>% 
  addMarkers(data = burn_icon, lng = ~x, lat = ~y,
             icon = list(
               iconUrl = burn_icon$image,
               iconSize = c(25, 25))) %>% 
  fitBounds(lng1 = burn_icon$x - 0.5,
            lat1 = burn_icon$y - 0.5,
            lng2 = burn_icon$x + 0.5,
            lat2 = burn_icon$y + 0.5) %>% 
  # Layers control
  addLayersControl(
    baseGroups = c("Daily Average", "Daily Maximum"),
    options = layersControlOptions(collapsed = FALSE)
  )
  
# remove smoke dispersion files from directory
unlink(paste(md_smoke,
                 "/smoke_dispersion_files", sep = ""), 
       recursive = TRUE)

```

```{r leaflet_map_show, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# show leaflet map
day_one_leaflet

``` 

## Meterological Conditions  

```{r firepoker_link, echo = FALSE, include = FALSE }

## Generate firepoker link

# get burn location
burn_loc <- burn_info %>% 
  dplyr::select(-name)

# specify coordinates
coordinates(burn_loc) <- c("lon", "lat")
proj4string(burn_loc) <- CRS("+init=epsg:4326")

# transform to bbox CRS
# burn_loc_3857 <- spTransform(burn_loc, CRS("+init=epsg:3857"))@coords

# bbox data: xmin, ymin, xmax, ymax
# bbox_ctds <- c(burn_loc_3857[1] + -1016307,
#                burn_loc_3857[2] - 489197,
#                burn_loc_3857[1] - -1016307,
#                burn_loc_3857[2] + 489197)



# read in dispersion breakpoints and id the state where burn is happening
disp_brkpts_df <- read_excel("../raw_data/disp_breakpoints.xlsx")
states <- st_read("../gis/states")
burn_info_sf <- st_as_sfc(burn_loc)
state_sf <- st_transform(states, crs = st_crs(burn_info_sf))
burn_state <- as.character(states$STATE_ABBR[st_intersects(burn_info_sf, state_sf)[[1]]])


disp_brkpts <- disp_brkpts_df %>% 
  filter(state == burn_state) 


# create firepoker link
  # create firepoker link
  fp_url <- paste("https://www.weather.gov/dlh/firepoker?lat=",
                  round(burn_loc$lat, 3),
                  "&lon=",
                  round(burn_loc$lon, 3),
                  "&clat=38.967&clon=-97.267&zoom=7.000&bbox=[-16540299.098,2518065.675,-5115162.882,6915872.719]&layers=USStates|ForecastDot|Domain|SurfaceFronts|Radar|&fwf=F&dispersion=",
                  paste(0, 
                        disp_brkpts[2], 
                        disp_brkpts[3],
                        disp_brkpts[4],
                        disp_brkpts[5],
                        sep = ","),
                  "&ndfd=WindGust&nohelp=t#",
                  sep = "")
```

The predicted weather data, including fire behavior and smoke parameters, for the area are shown below. They can also be found [here](`r fp_url`). 

```{r met_data, echo=FALSE, out.width="100%", out.height="100%",fig.show='hold',fig.align='center'}
# include screeshot of firepoker website
include_graphics(paste(md_smoke,
                       "/firepoker.PNG", sep = ""))

```














